---
title: "607 -Project 2 - Oscars"
author: "Jose Mawyin"
date: "10/6/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tibble)
library(tidyr)
library(dplyr)
library(data.table)
library(ggplot2)
```

## Oscar Winners Demographics

###How has ethnicity shaped the movie industry?
We will look at how the ethnic mix of Oscar winners has changed or nor in time by looking at a dataset that has collated Oscar Winner information from the years 1927 to 2014.
```{r}
Oscars <- read.csv("/Users/josemawyin/Downloads/Oscars-demographics-DFE.csv")
```

The dataset has many fields but of interest are race_ethnicity year_of_award
```{r}
str(Oscars)
dim(Oscars)
```
Let's first find how many Oscars have been awarded based on ethnicity:
```{r}
summary(Oscars$race_ethnicity)
```

Then let's find when was the first Oscar awarded to an ethnicy other than "White":
```{r}
count(Oscars, year_of_award, race_ethnicity) %>% filter(race_ethnicity != "White") %>%  arrange(year_of_award)
```
We can see that in the period of 1927 to 1940 no Oscars were awarded to non-White winners. 

Below we can see how the ethnic inclusivity has increased since the later quarter of the 20th century until the present.
```{r}
ggplot(Oscars, aes(year_of_award, fill = race_ethnicity)) +
  geom_histogram(binwidth = 5) + ylab("Total Oscars Awarded") + xlab("Years") + ggtitle("Oscar Winners by Ethnicity")
```

Lack of inclusion for unrepresented members of the movie community is evident in the data shown in this study. However, there is been a change in the number of Oscar's awarded to other-than-white  Further study could see how does the number of winners compared to the total number of movie industry workers broke down by ethnicity. Are other-than-white represented?